---
title: "Viz Template"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading Excel files
library(lubridate)  # for dates
library(readxl)     # for reading excel files

# packages for using custom fonts
# library(systemfonts)
# library(textshaping)
library(ragg)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC: <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer - <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts - <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric Scherer

```{r, eval=FALSE}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space) - Cédric Scherer

```{r, eval=FALSE}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r, eval=FALSE}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

Saving using custom fonts with ragg

```{r}
ggplot(mtcars, aes(hp, mpg)) +
  geom_point() +
  labs(title = "22222222222222222") +
  theme(text = element_text(family = "Roboto"),
        plot.title = element_text(size = 50))

ggsave("filename.png", device = ragg::agg_png())
```

## Colour palettes

## Remember it is an iterative process

Trying to recreate and enahnce figures from "T[he People and Nature Survey for England: Data and publications from Adults survey year 1 (April 2020 - March 2021) (Official Statistics) main findings](https://www.gov.uk/government/statistics/the-people-and-nature-survey-for-england-data-and-publications-from-adults-survey-year-1-april-2020-march-2021-official-statistics/the-people-and-nature-survey-for-england-data-and-publications-from-adults-survey-year-1-april-2020-march-2021-official-statistics-main-finding)"

## Import data

```{r}
# ../../../datasets/people_and_nature_survey/People_and_Nature_Q1_Q4.xlsx HOME
# ../../datasets/people_and_nature_survey/People_and_Nature_Q1_Q4.xlsx WORK

pan_raw <- readxl::read_xlsx("../../datasets/people_and_nature_survey/People_and_Nature_Q1_Q4.xlsx", sheet = "Data", guess_max = 26000)
```

## Clean data

```{r}
pan_clean <- pan_raw %>% 
  janitor::clean_names()
```

## Explore the data

```{r}

```

## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for comparisons?**

## Think about the audience

+-----------------------------------------------------------------------------------+-----------+
| To think about                                                                    | Notes     |
+===================================================================================+===========+
| Can you think of a group of people who might view the visualisations?             |           |
+-----------------------------------------------------------------------------------+-----------+
| Can you think of a specific person who might view the visualizations?             |           |
+-----------------------------------------------------------------------------------+-----------+
| What do want the audience to know or be able to after viewing the visualizations? |           |
+-----------------------------------------------------------------------------------+-----------+
| What expectations might the audience have about visualisations of this data?      |           |
+-----------------------------------------------------------------------------------+-----------+

## Ask some questions to be answered by the visualizations

1.  How does the proportion of people who have visited green space in the last 14 days change over the course of FY 2020/21?

2.  ...

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

+-----------------------------------------------------------------------------------------------------------------------------+---------------------+-----------------------------------------------+
| Question                                                                                                                    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============================================================================================================================+=====================+===============================================+
| 1.  How does the proportion of people who have visited green space in the last 14 days change over the course of FY 2020/21 |                     |                                               |
+-----------------------------------------------------------------------------------------------------------------------------+---------------------+-----------------------------------------------+

```{r}
a_y_14_fct_levels <- c("Other", "No visits","Any visits")

any_visits_by_month <- pan_clean %>% 
  
  mutate(any_visits_14 = if_else(any_visits_14 == "Dont know" | 
                                   any_visits_14 == "Prefer not to say",
                                 true = "Other", 
                                 false = any_visits_14)) %>% 
  
  select(wave, weight_percent, any_visits_14) %>%
  filter(!is.na(any_visits_14)) %>% 
  separate(wave, into = c("wave", "date"), sep = " - ") %>% 
  mutate(date = lubridate::my(date)) %>% 
  select(-wave) %>% 
  group_by(date, any_visits_14) %>% 
  summarise(n = sum(weight_percent)) %>% 
  mutate(month = format(date,"%B"), freq = n / sum(n),
         any_visits_14_fct = factor(any_visits_14,
                                    levels = a_y_14_fct_levels)) %>% 
  ungroup() 


ggplot(any_visits_by_month, aes(date, freq, fill = any_visits_14_fct)) +
  geom_area(alpha = 0.75) +
  
  scale_x_date(expand = c(0,10), date_breaks = "1 month",
               date_labels = "%b") +
  scale_y_continuous(expand = c(0,0),
                     labels = scales::percent_format(scale = 100)) +
  scale_fill_manual(values = c("white", "grey85", "#0EAB4E")) +
  
  labs(x = NULL) +
  theme_cjmr_explanatory() +
  theme(legend.position = "none")
  #facet_wrap(~any_visits_14_fct)

ggsave("viz_out/figure_1.svg", units = "mm", height = 70, width = 120)

# output for table
any_visits_by_month %>% 
  select(month, any_visits_14_fct, freq) %>% 
  pivot_wider(names_from = month, values_from = freq) %>% 
  arrange(desc(any_visits_14_fct)) %>% 
  write_csv("data_out/any_visits_by_month.csv")
  
```

**Calculate confidence intervals**

Just for the proportion of people who have visited in the last 14 days

```{r}
ci_plot_df <- any_visits_by_month %>% 

  # focus on proportion who have visted
  filter(any_visits_14_fct == "Any visits") %>% 
  
  # calculate standard error
  mutate(z = 1.96,
         se = sqrt(freq * (1 - freq) / n)) %>% 
  
  # calculate upper and lower bounds of 95% confidence interval
  mutate(lower_ci = freq - (z * se),
         upper_ci = freq + (z * se)) %>% 
  
  # prepare for plotting
  select(date, freq, lower_ci, upper_ci)

ggplot(ci_plot_df) +
  
  geom_ribbon(mapping = aes(x = date,
                            ymin = lower_ci,
                            ymax = upper_ci),
              fill = "#5CD58D", alpha = 0.5) +
  geom_line(mapping = aes(date, freq),
            colour = "#077C37") +
  
  # geom_errorbar(mapping = aes(x = date,
  #                           ymin = lower_ci,
  #                           ymax = upper_ci),
  #                colour = "#33BB6A", width = 5) +
  
  geom_point(mapping = aes(date, freq),
             colour = "#077C37") +
  
  scale_x_date(expand = c(0,10), date_breaks = "1 month",
               date_labels = "%b") +
  scale_y_continuous(expand = c(0,0),
                     labels = scales::percent_format(scale = 100,
                                                     accuracy = 5)) +
  labs(x = NULL) +
  
  theme_cjmr_explanatory()

ggsave("viz_out/figure_2.svg", units = "mm", height = 70, width = 120)
  

```

### Visualization x (question n)

| Question | Answer (i.e. title) | Notes to consider in design, chart types etc. |
|----------|---------------------|-----------------------------------------------|
|          |                     |                                               |

```{r}

```

### Visualization x (question n)

| Question | Answer (i.e. title) | Notes to consider in design, chart types etc. |
|----------|---------------------|-----------------------------------------------|
|          |                     |                                               |

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

| Purpose                 | Width | Height |
|-------------------------|-------|--------|
| Single plot for Twitter |       |        |
|                         |       |        |
|                         |       |        |

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a check-box.

+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Question to consider                                | Tips                                                        | Notes      |
+=====================================================+=============================================================+============+
| Are the **encodings** explained?                    | Air on the side of too much explanation ...                 |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Is the **context** explained?                       | Titles, subtitles                                           |            |
|                                                     |                                                             |            |
|                                                     | Annotate                                                    |            |
|                                                     |                                                             |            |
|                                                     | Use colour to highlight                                     |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Does it have **readability** like text?             | Think visual hierarchy *(as when laying out text)*          |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Do the **aesthetics** align with the topic/message? | Serious topic -> dark colours, straight-forward chart types |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+

## Remember it is an iterative process
