---
title: "Viz Template"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading Excel files
library(lubridate)  # for dates
library(readxl)     # for reading excel files

# packages for using custom fonts
# library(systemfonts)
# library(textshaping)
library(ragg)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC: <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer - <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts - <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric Scherer

```{r, eval=FALSE}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space) - Cédric Scherer

```{r, eval=FALSE}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r, eval=FALSE}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

Saving using custom fonts with ragg

```{r}
ggplot(mtcars, aes(hp, mpg)) +
  geom_point() +
  labs(title = "22222222222222222") +
  theme(text = element_text(family = "Roboto"),
        plot.title = element_text(size = 50))

ggsave("filename.png", device = ragg::agg_png())
```

## Colour palettes

## Remember it is an iterative process

Trying to recreate and enahnce figures from "T[he People and Nature Survey for England: Data and publications from Adults survey year 1 (April 2020 - March 2021) (Official Statistics) main findings](https://www.gov.uk/government/statistics/the-people-and-nature-survey-for-england-data-and-publications-from-adults-survey-year-1-april-2020-march-2021-official-statistics/the-people-and-nature-survey-for-england-data-and-publications-from-adults-survey-year-1-april-2020-march-2021-official-statistics-main-finding)"

## Import data

```{r}

```

## Clean data

```{r}

```

## Explore the data

```{r}

```

## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for comparisons?**

## Think about the audience

+-----------------------------------------------------------------------------------+----------+
| To think about                                                                    | Notes    |
+===================================================================================+==========+
| Can you think of a group of people who might view the visualisations?             |          |
+-----------------------------------------------------------------------------------+----------+
| Can you think of a specific person who might view the visualizations?             |          |
+-----------------------------------------------------------------------------------+----------+
| What do want the audience to know or be able to after viewing the visualizations? |          |
+-----------------------------------------------------------------------------------+----------+
| What expectations might the audience have about visualisations of this data?      |          |
+-----------------------------------------------------------------------------------+----------+

## Ask some questions to be answered by the visualizations

1.  ...

2.  ...

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

| Question | Answer (i.e. title) | Notes to consider in design, chart types etc. |
|----------|---------------------|-----------------------------------------------|
|          |                     |                                               |

```{r}

```

### Visualization x (question n)

| Question | Answer (i.e. title) | Notes to consider in design, chart types etc. |
|----------|---------------------|-----------------------------------------------|
|          |                     |                                               |

```{r}

```

### Visualization x (question n)

| Question | Answer (i.e. title) | Notes to consider in design, chart types etc. |
|----------|---------------------|-----------------------------------------------|
|          |                     |                                               |

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

| Purpose                 | Width | Height |
|-------------------------|-------|--------|
| Single plot for Twitter |       |        |
|                         |       |        |
|                         |       |        |

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a check-box.

+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Question to consider                                | Tips                                                        | Notes      |
+=====================================================+=============================================================+============+
| Are the **encodings** explained?                    | Air on the side of too much explanation ...                 |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Is the **context** explained?                       | Titles, subtitles                                           |            |
|                                                     |                                                             |            |
|                                                     | Annotate                                                    |            |
|                                                     |                                                             |            |
|                                                     | Use colour to highlight                                     |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Does it have **readability** like text?             | Think visual hierarchy *(as when laying out text)*          |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+
| Do the **aesthetics** align with the topic/message? | Serious topic -> dark colours, straight-forward chart types |            |
+-----------------------------------------------------+-------------------------------------------------------------+------------+

## Remember it is an iterative process
